---
title: "socioecon_analysis"
author: "KT Hobbs"
date: "15/06/2020"
output: html_document
---

```{r libs}
library(skimr)
```

```{r read}
df = read.csv(file = 'econ_kt.csv', stringsAsFactors = FALSE, na.strings="", header=TRUE)
colnames(df)
```


```{r }
df2 = df[,-c(1, 2, 25, 30:50,52)]

# summary(df2)
 skim(df2)

```


```{r}
df3 = na.omit(df2)

pc2 <- prcomp(df3)
summary(pc2)
```

```{r}
screeplot(pc2)
```


```{r}
pc2$rotation[,1:2]
```
PC1: higher comorbidities, more detached homes, less amenity dense, higher age of population, less higher education, poorer mental and physical health

PC2: higher cost of living, high rises, more unemployment/less in the work force, perceived high stress, more commuting altogether and more vehical transit usage, percieved good health, much more amenity dense? => *possibly indicative of a city*

With commute time, all variables load onto one component.

```{r}
library(dplyr)
df_na <- na.omit(df)
mutate(df3, total = df_na$TOTAL)
```



```{r}

suppressWarnings(biplot(pc2, xlabs = df_na$Location_x))
```

