---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

# Factor Analysis for PHU data

### Proximity Measures:

```{r packages, include = F}
suppressMessages(library(corrplot))
suppressPackageStartupMessages(library('cluster'))
suppressPackageStartupMessages(library('dplyr'))
```

```{r read and clean, include = F}
phu_orig <- read.csv('../data/PMD-en/PHU_FINAL_prop.csv')
phu <- phu_orig[,-c(1,2)]

# percent comorbidities clean ----
percents <- c('copd.percent', 'asthma.percent', 'smokers.percent', 'hbp.percent')
phu[,percents] <- lapply(phu[,percents], function(x) as.numeric(gsub("[\\%,]", "", x))/100)

# convert missing transit data to 0 ----
# (ASSUMPTION: there is no public transit in the corresponding PHU)
phu$prox_idx_transit[is.na(phu$prox_idx_transit)] <- 0


# PHU cols for pca ----
keep <- c('Location', 'asthma.percent', 'copd.percent', 'hbp.percent', 'smokers.percent', 
         #'FEMALEprop', 'MALEprop', 'TRANSGENDERprop','OTHERprop', 'UNKNOWNprop',
         #'CONTACTprop', 'NEITHERprop', 'TRAVEL.RELATEDprop',
         'prox_idx_emp', 'prox_idx_pharma', 'prox_idx_childcare', 'prox_idx_health',
         'prox_idx_grocery', 'prox_idx_educpri', 'prox_idx_educsec', 'prox_idx_lib',
         'prox_idx_parks', 'prox_idx_transit')

phu_prop <- subset(phu, select = keep)

#assign names phu df
rownames(phu_prop) <- phu_prop$Location

# grab only prox data
prox <- phu_prop[ , grepl( "prox" , names(phu_prop) ) ]

```

```{r prox corr, include = F}
cprox <- cor(prox)
corrplot(cprox, type = "upper", method = "circle", order = "FPC",
         main = "Proximity Correlations - First Principal Component Order")
```

Proximity to secondary education and libraries are generally uncorrelated with the rest.

**2-factor Factor Analysis to identify connectedness (or urban v rural):**

We can accept null hypothesis that factors are sufficient (1 factor is not, 3 is even less significant so 3 factors may be reasonable as well).

```{r FA proximity, include = F}
faprox2 <- factanal(prox[,-1], 2, scores="regression")
faprox2
```


Hierarchical clustering of factor analysis score output shows the same as on euclidean distance data.

```{r fa proximity cluster, include = F}
faclust <- diana(dist(faprox2$scores))
pltree(faclust, cex = 0.6, hang = -1, main = "Dendrogram of DIANA") 
```

### All PHU data

```{r phu corr, include = F}
cphu <- cor(phu_prop[,-1])
corrplot(cphu, type = "upper", method = "circle", order = "FPC",
         main = "PHU Data Correlations - First Principal Component Order")
```

The original PHU data is a singular invertible matrix. This means that rows or columns are linearly interdependent.

```{r FA PHU}
factanal(phu_prop[,-1], 2, scores="regression")
```

Assessing multicolinearity in the data on a linear model fit:

```{r lm multicolinearity}
tofit <- mutate(phu_prop, total = phu_orig$TOTALprop)
fit <- lm(total ~.-Location, data = tofit)

car::vif(fit)
```


["VIF quantifies the severity of multicollinearity in an ordinary least squares regression analysis"](https://en.wikipedia.org/wiki/Variance_inflation_factor)

