---
title: "R Notebook"
output:
  html_document:
    df_print: paged
  pdf_document: default
---
# COVID PHU-Level Clustering 

## Hierarchical Clustering Methods

```{r packages, include = F}
suppressPackageStartupMessages(library('dplyr'))
suppressPackageStartupMessages(library('tidyverse'))
suppressPackageStartupMessages(library('mclust'))
suppressPackageStartupMessages(library('factoextra'))
suppressPackageStartupMessages(library('cluster'))
suppressPackageStartupMessages(library('ggplot2'))
suppressPackageStartupMessages(library('dendextend'))
```

```{r read and clean data, include = F}
phu <- read.csv('../data/econ_kt.csv')

# remove duplicate PHUS
phu <- phu[-c(12,27),]

#assign names phu df
rownames(phu) <- phu$Location_x

# remove unneccessary cols
phu2 <- phu[,-c(1, 2, 25, 30:50,52, 75,76)]

# create df focussing on health and amenity measures (no socioecon data)
health_amen <- phu2[-c(1:26)]

```

### Agglomerative Clustering:

Data is first scaled then the euclidean distance is computed between each PHU. Clusters are formed based on least dissimilarities between observations.

*Group average clustering:* attempts to produce relatively compact that are relatively far apart.

*Complete clustering:* all union observations for two groups are least dissimilar - can produce clusters that violate "closeness" property (one observation may be more similar to observations in other groups).

*Single clustering:* grouped based on one dissimilarity measure common between two observations being very small - prone to chaining.

Single linkage appears to have chaining while average linkage groups are not as easily discernable. Complete linkage was chosen for further analysis.

```{r scale and dist, include = F}
#scale
phu_sc <- scale(health_amen)

#euc distance between each PHU is ok because all var are quantitative
phu_dist <- dist(phu_sc)

#clust
comp <- hclust(phu_dist, method = "complete")
av <- hclust(phu_dist, method = "average")
sing <- hclust(phu_dist, method = 'single')

plot(comp, main = "Complete Linkage") # best, suggests 2 groups
plot(av, main = "Average Linkage") # chaining
plot(sing, main = "Single Linkage") # chaining


```

```{r complete clust, echo = F}
# comp clust better visualization ----
comp_dend <- as.dendrogram(comp)
comp_col_dend <- color_branches(comp_dend, h = 3)
pdf('../visualizations/Dendrogram.pdf')
plot(comp_col_dend, main = "Complete Linkage")
dev.off()
```

Using 2 clusters...

```{r comp clust k 2, echo = F}

# 8 clusters according to previous plot ----
phu_cl <- mutate(health_amen, cluster = cutree(comp, k = 2)) 
count(phu_cl,cluster)

# VIS: cluster groups ----
pdf('Clusters.pdf')
fviz_cluster(list(data = phu_cl, cluster = cutree(comp, k = 2)), 
             main = "Cluster Plot k = 2 - Agglomerative Complete Hierarchical Clustering")
dev.off()

```

The agglomerative coefficient for complete linkage:

```{r ac comp, echo = F}
paste("Agglomerative Coefficient: ", coef.hclust(comp))
```

```{r k 3, echo = F}
# VIS: bar plots ----
phu_cl <- mutate(phu_cl, CovidCase = phu$TOTAL_prop) 
phu_cl <- mutate(phu_cl, Reporting_PHU = phu$Location_x) 
phu_cl <- mutate(phu_cl, Cluster = factor(phu_cl$cluster)) 

# clean location names
phu_cl$abbrev <- phu_cl$Reporting_PHU
phu_cl$abbrev <- str_remove(phu_cl$Reporting_PHU, "Health Unit")
# phu_amen_cl$abbrev <- str_replace(phu_cl$abbrev, "North Bay Parry Sound District", "North Bay/Parry Sound")
#phu_amen_cl$abbrev <- str_replace(phu_amen_cl$abbrev, "Haliburton, Kawartha, Pine Ridge", "HKPR")
#phu_amen_cl$abbrev <- str_replace(phu_amen_cl$abbrev, "Leeds, Grenville and Lanark District", "LGL")
phu_cl$abbrev <- str_replace(phu_cl$abbrev, "Kingston, Frontenac and Lennox & Addington Public Health", "KFL&A")
phu_cl$abbrev <- str_remove(phu_cl$abbrev, "District")
phu_cl$abbrev <- str_remove(phu_cl$abbrev, "County")
phu_cl$abbrev <- str_remove(phu_cl$abbrev, "The")
phu_cl$abbrev <- str_remove(phu_cl$abbrev, "of")
phu_cl$abbrev <- str_remove(phu_cl$abbrev, "City")
phu_cl$abbrev <- str_remove(phu_cl$abbrev, "Regional")
phu_cl$abbrev <- str_remove(phu_cl$abbrev, "Region")
phu_cl$abbrev <- str_remove(phu_cl$abbrev, "Counties")
phu_cl$abbrev <- str_remove(phu_cl$abbrev, "and")
phu_cl$abbrev <- str_remove(phu_cl$abbrev, "-")
phu_cl$abbrev <- str_remove(phu_cl$abbrev, "Area")
phu_cl$abbrev <- str_replace(phu_cl$abbrev, "  ", " ")
phu_cl$abbrev <- str_replace(phu_cl$abbrev, ",", "")
phu_cl$abbrev <- str_replace(phu_cl$abbrev, "Public", "")
phu_cl$abbrev <- str_replace(phu_cl$abbrev, "Department", "")
phu_cl$abbrev <- str_replace(phu_cl$abbrev, "Health", "")
phu_cl$abbrev <- str_replace(phu_cl$abbrev, "Services", "")
phu_cl$abbrev <- str_replace(phu_cl$abbrev, "Sudbury &", "Sudbury")

pdf('../visualizations/ClusterBar.pdf')
ggplot(phu_cl, aes(x=fct_reorder(abbrev,CovidCase, .desc=F), y = CovidCase, fill = Cluster)) + 
  geom_bar(stat="identity")  +
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  labs(x = "", y = "") +
  ggtitle('Proportion of COVID-19 Cases per Public Health Unit')
dev.off()

```


```{r}

colnames(phu_cl)

ggplot(phu_cl, aes(x=fct_reorder(abbrev,amenity_dense, .desc=F), y = amenity_dense, fill = Cluster)) + 
  geom_bar(stat="identity")  +
  theme(axis.text.x=element_text(angle=45,hjust=1)) +
  labs(x = "", y = "") +
  ggtitle('Amenity Density per Public Health Unit')
```

### Divisive Hierarchical Clustering - DIANA
All observations begin as a single cluster and are divisible separated based on largest average dissimilarity between an observation and all others.

Hierarchical clustering will always cluster even if there is no grouping in the data. The divisive coefficient is on a 0-1 scale, representing the extent to which the hierarchical structure produced by a dendrogram actually represents the data (closer to 1 = better representation).

```{r DIANA, echo = F}
hc <- diana(phu_dist)
paste("Divisive Coefficient: ", hc$dc)
```

```{r DIANA VIS, echo = F}
pltree(hc, cex = 0.6, hang = -1, main = "Dendrogram of DIANA")
clust <- cutree(hc, k = 4)
fviz_cluster(list(data = phu_dist, cluster = clust), main = "Cluster Plot DIANA - Divisive Hierarchical Clustering")

```

Covid Case Total Proportions (per PHU) data and DIANA cluster assignment are added to `phu_prop` to assess groupings by covid output.

```{r diana bar, echo = F}
# VIS: bar plots ----
phu_cl <- mutate(health_amen, CovidCaseStrat = phu$TOTAL_prop) 
phu_cl <- mutate(phu_cl, cluster = clust) 
phu_cl <- mutate(phu_cl, Reporting_PHU = phu$Location_x) 

ggplot(phu_cl, aes(x=fct_reorder(Reporting_PHU,CovidCaseStrat, .desc=F), y = CovidCaseStrat, fill = factor(cluster))) + 
  geom_bar(stat="identity")  +
  theme(axis.text.x=element_text(angle=90,hjust=1)) +
  labs(x = "PHU", y = "Proportion of PHU with Covid Cases")

```


## Mixture Model Clustering Methods 

### EM Algorithm

Soft version of K-means clustering that is based on probabilities rather than deterministic assignments of points to groups. Used package `Mclust`.

For all parameters:

```{r EM all params, include = F}
# EM by all params ----
mphu <-Mclust(health_amen, G=1:3)
plot(mphu, what="classification")
```


Between PHU and Proportion of Cases:

```{r EM cases, include = F}
# EM by total cases ----

mcovid <- Mclust(phu$TOTAL_prop, G=1:2)
plot(mcovid, what="classification")
```

```{r EM case classes}
# EM view clasess ----
classes <- mutate(as.data.frame(mcovid$classification), loc = phu$Location_x)
colnames(classes)[1] <- 'class'
classes <- classes[order(classes),]
classes
```


### Multidimensional Scaling

Uses euclidean distances of scaled PHU data. Similarly represents hierarchical clustering (Toronto is its own cluster, denser regions such as Peel, Halton, York, Waterloo, Ottawa...etc. separate from smaller districts).

```{r}
plot(cmdscale(phu_dist)[,1], cmdscale(phu_dist)[,2], type = "n", 
     xlab = "", ylab = "", 
     asp = 1, axes = FALSE, 
     main = "MDS Results on phu_dist")
text(cmdscale(phu_dist)[,1], cmdscale(phu_dist)[,2], rownames(cmdscale(phu_dist)), cex = 0.6)


```
