---
title: "ltc_analysis"
author: "NL"
date: "15/06/2020"
output: html_document
---

## Import and prepare the data
```{r results="hide"}
library(skimr)

# Import the data
df <- read.csv("merged_LTC_odhf_quality.csv", na.strings="", header=TRUE)
skim(df)
```


```{r, include=FALSE}
# Add derived variables
# Add a column with outbreak status
df$outbreak <- with(df, ifelse(is.na(status), 'no', 'yes'))

# Add a column of LHIN health regions
unique(df[['LHIN']])
df$region[df$LHIN=='North West' | df$LHIN=='North East'] <- "North"
df$region[df$LHIN=='Toronto Central'] <- "Toronto"
df$region[df$LHIN=='Champlain' | df$LHIN=='Central East' | df$LHIN=='South East'] <- "East"
df$region[df$LHIN=='Erie St. Clair' | df$LHIN=='Hamilton Niagara Haldimand Brant (Hnhb)' | df$LHIN=='South West' | df$LHIN=='Waterloo Wellington'] <- "West"
df$region[df$LHIN=='Mississauga Halton' | df$LHIN=='Central West' | df$LHIN=='Central' | df$LHIN=='North Simcoe Muskoka'] <- "Central"

# Sum of complaints and critical
df$total_cc <- df$total_complaints + df$total_critical
df$X5y_cc <- df$X5y_complaints + df$X5y_critical
df$X2y_cc <- df$X2y_complaints + df$X2y_critical

# Sum of complaints, critical and withOrders
df$total_ccw <- df$total_critical + df$total_complaints + df$total_withOrders
df$X5y_ccw <- df$X5y_complaints + df$X5y_critical + df$X5y_withOrders
df$X2y_ccw <- df$X2y_complaints + df$X2y_critical + df$X2y_withOrders

# Number of non-complaints inspections
df$total_noncomplaints <- df$total_inspections - df$total_complaints
df$X5y_noncomplaints <- df$X5y_inspections - df$X5y_complaints
df$X2y_noncomplaints <- df$X2y_inspections - df$X2y_complaints

```

```{r, echo=FALSE}
# Change variable types
# Update data types for some columns
df[, 'outbreak'] <- as.factor(df[, 'outbreak'])
df[, 'region'] <- as.factor(df[, 'region'])

# Change data type of character
listofcol = list('name', 'cleaned_name', 'address', 'city', 'postal_code', 'CSDname')
for (each in listofcol){
  df[, each] <- as.character(df[, each])
}

skim(df)
```


```{r, include=FALSE}
# Generate the data for analysis
# Create lists of variables used to filter data for analysis
keep = c("outbreak", "home_type", "number_beds")
profile = c("short_stay", "residents_council", "family_council", "accreditation")
location = c('address', 'LHIN', 'region', 'city', 'postal_code', 'CSDname', 'CSDuid', 'latitude', 'longitude')
quality = c("antipsychotic_percent", "depression_percent", "falls_percent", "pressure_ulcers_percent", "pain_percent")

all_inspections = c("total_inspections", "X5y_inspections", "X2y_inspections")
complaints = c("total_complaints", "X5y_complaints", "X2y_complaints")
critical = c("total_critical", "X5y_critical", "X2y_critical") 
noncomplaints = c("total_noncomplaints", "X5y_noncomplaints", "X2y_noncomplaints")
withOrders = c("total_withOrders", "X5y_withOrders", "X2y_withOrders")
cc_ccw = c("total_cc", "X5y_cc", "X2y_cc", "total_ccw", "X5y_ccw", "X2y_ccw")

covid = c("confirmed_resident_cases", "resident_deaths", "confirmed_staff_cases")

data <- subset(df, select = c(keep, profile, 'region', quality, all_inspections, complaints, critical, noncomplaints, withOrders, cc_ccw))

```


```{r, include=FALSE}
## Assessing skewness
# Check for R skewness in numeric predictors
library(e1071)

listofcols = c('number_beds', quality, all_inspections, complaints, noncomplaints, critical, withOrders, cc_ccw)

for (each in listofcols){
  print(each)
  print(skewness(data[[each]]))
  qqnorm(data[[each]], main = each)
  qqline(data[[each]], col = 'red')
}

```

We see above that the inspections and number of beds data are highly right skewed with skewness values > 1. The quality data are not as right skewed with skewness values between 0 and 1, except for pain_percent which has a skewness value of 2.5. 

The highly skewed variables will have to be transformed for the analysis, but first we explore the variables to assess for any zero values. 


```{r, include=FALSE}
for (each in listofcols){
  print(each)
  print(summary(df[[each]]))
}
```
We see that the following variables contain zeros:
  - antipsychotic_percent
  - pressure_ulcers_percent
  - pain_percent
  - complaints group
  - critical group
  - withOrders group
  - cc group
  - X2y_noncomplaints
  - X2y_ccw

## Explore the variables for transformation
```{r, fig.show="hide"}
# Create a list of numeric variables with 0 values
list0s <- c('antipsychotic_percent', 'pressure_ulcers_percent', 'pain_percent', complaints, 'X2y_noncomplaints', critical, withOrders, 'total_cc', 'X5y_cc', 'X2y_cc', 'X2y_ccw')

library(rcompanion)
listofcols <- setdiff(listofcols, list0s)

# Visualize the effect of square root and log transformations on the data without 0s
for (each in listofcols) {  
  plotNormalHistogram(x = data[[each]], main = each)
  plotNormalHistogram(x = sqrt(data[[each]]), main = c(each, 'sqrt trans'))
  plotNormalHistogram(x = log(data[[each]]), main = c(each, 'log trans'))
}

```

A square root transformation seems best for the following data:
  - number_beds
  
A log transformation seems better for the following data:
  - all_inspections group
  - total_noncomplaints
  - X5y_noncomplaints
  - total_ccw
  - X5y_ccw
  
```{r, fig.show="hide"}
# Visualize the effect of a square root transformation on the variables with 0s
for (each in list0s) {  
  plotNormalHistogram(x = data[[each]], main = each)
  plotNormalHistogram(x = sqrt(data[[each]]), main = c(each, 'sqrt trans'))
}

```
As shown, in all cases where variables contain 0s, a square root transformation seems to improve the distribution. 

## Transform the data
```{r }
# Log transform in place
data[c(all_inspections, 'total_noncomplaints', 'X5y_noncomplaints', 'total_ccw', 'X5y_ccw')] <- log(data[c(all_inspections, 'total_noncomplaints', 'X5y_noncomplaints', 'total_ccw', 'X5y_ccw')])

# Square root transform in place
data[c('number_beds', list0s)] <- sqrt(data[c('number_beds', list0s)])

skim(data)
```


```{r }
# Inspect the distribution of the categorical variables
xtabs(~outbreak + home_type, data=df)
xtabs(~accreditation + home_type, data=df)
xtabs(~short_stay + home_type, data=df)
xtabs(~family_council + home_type, data=df)
xtabs(~residents_council + home_type, data=df)
xtabs(~residents_council + outbreak, data=df)
xtabs(~accreditation + outbreak, data=df)
xtabs(~short_stay + outbreak, data=df)
xtabs(~family_council + outbreak, data=df)
xtabs(~family_council + residents_council, data=df)
```

## Fit logistic regression
```{r }
# Fit logistic regression using transformed data
fit = glm(outbreak~., family=binomial, data=data)
summary(fit)
```

### Backwards selection
```{r, include=FALSE}
fit <- update(fit, .~. -pressure_ulcers_percent)
summary(fit)
fit <- update(fit, .~. -X5y_complaints)
summary(fit)
fit <- update(fit, .~. -total_critical)
summary(fit)
fit <- update(fit, .~. -total_cc)
summary(fit)
fit <- update(fit, .~. -X2y_withOrders)
summary(fit)
fit <- update(fit, .~. -depression_percent)
summary(fit)
fit <- update(fit, .~. -falls_percent)
summary(fit)
fit <- update(fit, .~. -X2y_complaints)
summary(fit)
fit <- update(fit, .~. -residents_council)
summary(fit)
fit <- update(fit, .~. -short_stay)
summary(fit)
fit <- update(fit, .~. -X5y_ccw)
summary(fit)
fit <- update(fit, .~. -total_ccw)
summary(fit)
fit <- update(fit, .~. -total_withOrders)
summary(fit)
fit <- update(fit, .~. -X5y_cc)
summary(fit)
fit <- update(fit, .~. -X2y_inspections)
summary(fit)
fit <- update(fit, .~. -accreditation)
summary(fit)
fit <- update(fit, .~. -X5y_inspections)
summary(fit)
fit <- update(fit, .~. -antipsychotic_percent)
summary(fit)
fit <- update(fit, .~. -pain_percent)
summary(fit)
fit <- update(fit, .~. -X2y_noncomplaints)
summary(fit)
fit <- update(fit, .~. -X2y_ccw)
summary(fit)
fit <- update(fit, .~. -X2y_cc)
summary(fit)
fit <- update(fit, .~. -X2y_critical)
summary(fit)
fit <- update(fit, .~. -X5y_critical)
summary(fit)
fit <- update(fit, .~. -X5y_noncomplaints)
summary(fit)
fit <- update(fit, .~. -total_noncomplaints)
summary(fit)
fit <- update(fit, .~. -X5y_withOrders)
summary(fit)
fit <- update(fit, .~. -total_inspections)
summary(fit)
fit <- update(fit, .~. -home_type)

```


```{r }
# Backwards selection
summary(fit)
```

As shown when all variables are included in the initial regression and then backwards selection is used for variable selection, the significant variables are number of beds, total complaints and region. 

```{r }
# Calculate McFadden's pseudo R2
ll.null <- fit$null.deviance/-2
ll.proposed <- fit$deviance/-2
print((ll.null - ll.proposed)/ll.null)
1 - pchisq(2*(ll.proposed - ll.null), df = (length(fit$coefficients)-1)) #pvalue

```

## Visualize the model
```{r, echo=FALSE}
# Create a new dataframe showing probability of outbreak and outbreak status
predicted.data <- data.frame(probability.of.outbreak=fit$fitted.values, outbreak=data$outbreak)

# Sort the above df
predicted.data <- predicted.data[order(predicted.data$probability.of.outbreak, decreasing=FALSE),]

# Add new col to df with rank
predicted.data$rank <- 1:nrow(predicted.data)

library(ggplot2)
library(cowplot)

ggplot(data = predicted.data, aes(x=rank, y=probability.of.outbreak)) +
  geom_point(aes(color=outbreak), alpha = 0.5, shape = 4, stroke = 1) +
  xlab("Index") + ylab("Predicted probability of an outbreak") +
  ggtitle("Outbreak Status Ordered By Predicted Probability of an Outbreak") + 
  scale_color_manual(values = c('blue', 'red'))

```


```{r }
# Repeat the logistic regression with selected variables
data_selected <- subset(df, select= c('outbreak', 'home_type', 'number_beds', noncomplaints, complaints, withOrders))
skim(data_selected)
fit = glm(outbreak~., family=binomial, data=data_selected)
summary(fit)
```

```{r, include=FALSE}
fit <- update(fit, .~. -X2y_withOrders)
summary(fit)
fit <- update(fit, .~. -X2y_complaints)
summary(fit)
fit <- update(fit, .~. -X2y_noncomplaints)
summary(fit)
fit <- update(fit, .~. -total_withOrders)
summary(fit)
fit <- update(fit, .~. -X5y_noncomplaints)
summary(fit)
fit <- update(fit, .~. -X5y_complaints)
summary(fit)
fit <- update(fit, .~. -X5y_withOrders)

```


```{r }
# Backwards Selection
summary(fit)
```


```{r }
### Repeat the logistic regression with selected variables
data_selected <- subset(df, select= c('outbreak', 'home_type', 'number_beds', all_inspections, complaints, withOrders))
skim(data_selected)
fit = glm(outbreak~., family=binomial, data=data_selected)
summary(fit)

```


```{r, include=FALSE}
fit <- update(fit, .~. -X2y_complaints)
summary(fit)
fit <- update(fit, .~. -X2y_withOrders)
summary(fit)
fit <- update(fit, .~. -X2y_inspections)
summary(fit)
fit <- update(fit, .~. -total_withOrders)
summary(fit)
fit <- update(fit, .~. -X5y_inspections)
summary(fit)
fit <- update(fit, .~. -X5y_complaints)
summary(fit)
fit <- update(fit, .~. -X5y_withOrders)
summary(fit)

```
```{r }
# Backwards Selection
summary(fit)
```

```{r }

```


```{r }

```

